---
title: Census 2016 packages
author: Hugh Parsonage
date: '2017-08-31'
slug: census-2016-packages
categories: ["R"]
tags: ["R"]
---

I've released two packages to access summary tables provided by the ABS.
The first `Census2016` is on CRAN and offers a limited set of data to comply with CRAN requirements.

A more substantial release is provided by the package `Census2016.DataPack` which contains hundreds of tables from the ABS Census 2016 data packs.

## `Census2016.DataPack`

### Motivation

The obvious motivation is that the 2016 census data packs hold a wealth of useful information that will be useful for researchers for years.
Bundling this data into a package improves the reproducibility of research,  as well as making it easier for researchers to *quickly* start analysis.

But I think the best way to explain my motivation for producing this package is to show you a variable name from Table G38 of the data pack:

> `Se_d_r_or_t_h_t_Tot_NofB_0_ib`

There are two problems with the data packs:

1. The variable names are arcane.
2. The data is not tidy: subtotals and subvariables lurk among the variable names.

In summary, the goals of `Census2016.DataPack` are:

1. To tidy the data so that the tables are normalized.
2. To provide at all costs readable variable names.
3. To use predictable table names and structure to support autocompletion.

### Installation

The package is not available on CRAN as it is far too large. You can install the package by running

```{r Installation}
if (!requireNamespace("Census2016.DataPack", quietly = TRUE)) {
  install.packages("Census2016.DataPack", 
                   repos = "https://hughparsonage.github.io/drat/", 
                   type = "source")
}
```

and attaching it in the usual way

```{r attachCensus2016DataPack}
library(Census2016.DataPack)
```

I'll also attach some other packages for analysis and prettiness.

```{r loadPackages}
library(magrittr)
library(data.table)
library(hutils)
```


### Usage 

As described in the motivation, all the tables have predictable, easy-to-read table names.
The drawback is that some of their names are very long.
This is unfortunate but in my view well worth it: it's trivial to adopt your own name for your work, but far more difficult to misread or be confused by a table.

For example, the number of persons in each `SA3` by age and sex is just 

```{r sa1-age-sex}
SA3__Age_Sex %>%
  data.kable
```

The real advantage of this package comes when using RStudio. 
Generally, you know what geographical area you need data for.
With that, you just start typing the three-letter code for the area (`SA3` in this case), 
wait for RStudio to access the internal data base, then select the most appropriate table that appears in the drop-down menu.

## Extensions

### New data in later releases

New data in the November release will be included in a later release.

### Select particular variables out-of-order

* Something like `SELECT(Sex, HoursHousekeeping, Age) %FROM% SA2` as a promise. 
* `see_question` like in `Census2016` (currently including this would endanger the 100 MB limit).
